Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS

### 2025-04-26 report/draft.qmd
update

China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy. This period yielded high average GDP growth rates alongside considerable improvements in poverty reduction, health, and education [@world_bank_world_2024; @ravallion_chinas_2009]. This growth, however, manifested unevenly across time, geography, and economic sectors. Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016]. Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024]. China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].

Parallel to this economic expansion, China experienced a substantial increase in wealth and income inequality. Analysis combining diverse data sources reveals a marked shift in distribution between 1978 and 2015; the wealth share controlled by the top 10% increased from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished considerably [@piketty_capital_2019]. Wealth inequality escalated rapidly, with the Gini coefficient for household wealth per capita rising markedly between 2002 and 2013, reaching high levels by international comparison [@knight_why_2022; @tan_earnings_2017]. Some research suggests that overall income inequality may have reached a peak around 2008-2010, potentially followed by a gradual decline or stabilization, linked partly to narrowing structural divides [@zhang_survey_2021; @kanbur_great_2021]. It is well-documented that China's inequality dynamics intersect closely with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021]. The urban-rural income ratio remained high even in the late 2010s, despite some narrowing from its 2009 peak, and regional income differences persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].

This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS). The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income. The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality. A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions. Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups. This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.

@@

Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62. A key finding is the role of housing wealth, which grew from 53% to 73% of total household assets and accounted for an increasing share of wealth inequality, rising from 64% to 79% of the total. The authors attribute the intensification of wealth inequality during these years to a combination of factors, including differential saving rates favoring the wealthy, and uneven house price inflation which benefited homeowners in major urban centers.

These trends in income and wealth inequality are deeply intertwined with persistent structural divides within China. Significant rural-urban and regional disparities remain central to understanding China's inequality landscape [@zhang_survey_2021; @xie_income_2014]. The urban-rural income ratio, despite some decline after peaking around 3.3 in 2009, remained elevated at approximately 2.7 in 2019 [@zhang_survey_2021]. This gap is reinforced by the *hukou* (household registration) system, which limits rural migrants' access to urban services and opportunities, even as migration increased substantially from 1990 to 2019. Regional inequality also persists, with coastal areas generally outpacing inland regions [@liao_sixty_2016]. While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@zhang_survey_2021; @kanbur_great_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.

@@

Furthermore, the pattern of economic growth influenced distributional outcomes. Research indicates that growth originating in the agricultural and rural sectors was more effective at reducing both poverty and overall inequality compared to growth in the industrial or service sectors [@ravallion_chinas_2009]. This suggests the importance of the initial economic structure and the specific mechanisms linking growth to income gains across population segments. Conversely, the rise in inequality itself appeared to impede poverty reduction; provinces that began the reform period with higher levels of inequality experienced slower subsequent progress in reducing poverty, partly because growth was less effective at lifting incomes at the lower end of the distribution in these areas [@ravallion_chinas_2009]. Specific policies, such as the initial agrarian reforms and adjustments to agricultural procurement prices, also demonstrably affected poverty and inequality levels. While income inequality may have plateaued or slightly decreased post-2010, linked to narrowing structural gaps and potentially policy interventions [@zhang_survey_2021; @liao_sixty_2016], wealth concentration, influenced by real estate appreciation, continued its upward trend well into the 2010s [@knight_why_2022; @piketty_capital_2019].

@@

Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels. As with most survey data, CHFS figures are subject to potential limitations. Asset and income data rely on self-reporting, which can be affected by reporting error or bias. Respondent privacy is protected through methods including the top-coding of certain high-value income and asset variables; this may result in an underestimation of the upper tails of the respective distributions and potentially impact inequality metrics. Researchers should also note that minor adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019, may require consideration for longitudinal comparisons. Missing data are handled through specific coding and imputation for certain variables.

The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis. The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior. Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others [@tan_earnings_2017; @gan_report_2014; @liu_health_2022].

@@

# Analysis
## Summary Statistics
::: {#tbl-tbl1}

@@

Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.

@@

Compared to the 2011 data, the Gini coefficient for income has decresed from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019. The Gini coefficient for wealth has decreased from 0.761 in 2011 to 0.701 in 2017 and continued to decrease to 0.667 in 2019. The results show that China's income and wealth inequality has been decreasing since 2011. The finding is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021, that China's inequality peaked around 2010 and started to decrease since then.

Notably, while the growth rate of mean and median income from 2011 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, where the growth of mean and median wealth greatly surpassed income growth. This later-period divergence may reflect factors such as significant asset price appreciation relative to income gains.

## Factor decomposition of wealth inequality

To understand the sources driving wealth inequality in China, this section applies the decomposition method developed by @lerman_income_1985. According to this method, the contribution of each wealth component to the overall Gini coefficient is decomposed as follows:

@@

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking [@lerman_income_1985].

@@

Summary Statistics of Wealth, 2017 and 2019
:::

@tbl-wealth-summary summarizes household wealth allocation in China for 2017 and 2019. Housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types. Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets. Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.

Financial equity notably increased from 2017 to 2019, with mean values rising substantially from 103,355 RMB to 161,199 RMB and median values more than doubling from 17,170 RMB to 39,251 RMB. Correspondingly, financial assets increased their share in total household wealth from around 12% in 2017 to approximately 15% in 2019. Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate]. However, later results in this paper show that the gini coefficient for land in the rural area was 89.7% in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas. This shows that dispite a wide distribution of land ownership, the distribution of land is highly unequal in modern China.

@@

Source: Author's calculation based on the 2017 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2017
:::

@@

Factor Decomposition of Wealth Gini Coefficient, 2019
:::

@tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively. Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality (72.7% in 2017, 72.3% in 2019). This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011. Financial equity emerged as another important factor, increasing its contribution from approximately 11.5% in 2017 to about 13.5% in 2019. In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.

Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019. Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors. To explore heterogeneity in these patterns, the following sections decompose wealth inequality within specific demographic subgroups.

>>>>

China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy. This period yielded high average GDP growth rates averaged over 9% per year along with considerable improvements in poverty reduction, health, and education [@world_bank_world_2024]. This growth, however, manifested unevenly across time, geography, and economic sectors. Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016]. Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024]. China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].

Parallel to this economic expansion, China experienced a significant rise in wealth and income inequality. @piketty_capital_2019 reveals a marked shift in wealth distribution between 1978 and 2015, highlighting that the wealth share controlled by the top 10% increased substantially, from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished significantly. This trend of rising inequality is further corroborated by the Gini coefficient for household wealth per capita, which increased notably from 0.5 in 2002 to 0.63 in 2013, according to an official national survey [@knight_why_2022]. Such levels of inequality are notably high by international standards [@tan_earnings_2017]. Furthermore, China's inequality dynamics are closely intertwined with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average income and wealth levels compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021]. Despite some narrowing since its peak in 2009, the urban-rural income ratio remained high in the late 2010s, and significant regional income disparities persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].

This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS). The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income [@kanbur_great_2021; @zhang_survey_2021]. The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality. A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions. Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups. This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.

@@

Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality using the Chinese Household Income Project (CHIP) data, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62. A significant contributor to this increase was housing wealth, which expanded from 53% to 73% of total household assets. Housing also became increasingly central to wealth inequality, with its contribution rising from 64% to 79% of the overall inequality. The authors attribute the intensification of wealth inequality during these years to factors such as differential saving rates favoring the wealthy and uneven house price inflation that particularly benefited homeowners in major urban centers.

These trends in income and wealth inequality are deeply intertwined with persistent structural divides within China. Significant rural-urban and regional disparities remain central to understanding China's inequality landscape [@zhang_survey_2021; @xie_income_2014]. The urban-rural income ratio, despite some decline after peaking around 3.3 in 2009, remained elevated at approximately 2.7 in 2019 [@zhang_survey_2021]. This gap is reinforced by the *hukou* (household registration) system, which limits rural migrants' access to urban services and opportunities, even as migration increased substantially from 1990 to 2019. Regional inequality also persists, with coastal areas generally outpacing inland regions [@liao_sixty_2016]. While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@kanbur_great_2021; @zhang_survey_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.

@@

@@

Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels [@gan_report_2014]. As with most survey data, CHFS figures are subject to potential limitations. Asset and income data rely on self-reporting, which can be affected by reporting error or bias. Respondent privacy is protected through methods including the top-coding of certain high-value income and asset variables; this may result in an underestimation of the upper tails of the respective distributions and potentially impact inequality metrics. It is also worth noting that potential adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019 which was not present in 2017, may require consideration for longitudinal comparisons.

The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis. The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior. Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others.

@@

# Analysis
## Summary Statistics
This section provides an overview of household income and wealth in China from 2011 to 2019, demonstrating that the 2017 and 2019 waves of the CHFS align with earlier studies in confirming a declining trend in China's inequality since 2010.

::: {#tbl-tbl1}

@@

*All figures are reported in RMB at constant 2017 prices.*

*Sources: @tan_earnings_2017's calculation based on the 2010 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.*

@@

Compared to 2010 data, the Gini coefficient for income decreased from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019. Similarly, the Gini coefficient for wealth declined from 0.761 in 2011 to 0.701 in 2017, and continued to decrease to 0.667 in 2019. These results indicate a consistent reduction in China's income and wealth inequality since 2010, aligning with findings from @kanbur_great_2021 and @zhang_survey_2021, which also report that China's inequality peaked around 2010 before starting to decline.

Notably, while the growth rate of mean and median income from 2010 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, when the growth in mean and median wealth considerably surpassed that of income. This later-period divergence may reflect factors such as significant asset price appreciation relative to income growth.

## Factor decomposition of wealth inequality
This section employs the decomposition method developed by @lerman_income_1985 to identify the specific contributions of different asset types to overall wealth inequality in China. The decomposition facilitates a detailed examination of how various asset components individually contribute to the aggregate wealth inequality measured by the Gini coefficient. This approach enables an analysis of the structural composition of inequality, clarifying not only the relative magnitude of each asset type's impact but also the extent to which each correlates with the overall wealth distribution. According to this method, the overall wealth Gini coefficient can be decomposed into contributions from each asset component as follows:

@@

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking. The summary statistics of each asset type are presented in @tbl-wealth-summary.

@@

*Source: Author's calculation based on the 2017 and 2019 CHFS data.*

*Summary Statistics of Wealth, 2017 and 2019*
:::

The data shows that housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types. Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets. Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.

Financial equity notably increased from 2017 to 2019, with mean values rising substantially from 103,355 RMB to 161,199 RMB and median values more than doubling from 17,170 RMB to 39,251 RMB. Correspondingly, financial assets increased their share in total household wealth from around 12% in 2017 to approximately 15% in 2019. Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate]. However, later results in this paper show that the gini coefficient for land in the rural area was as high as 0.9 in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas. This shows that dispite a wide distribution of land ownership, the distribution of land is highly unequal in modern China.

@@

*Source: Author's calculation based on the 2017 CHFS data.*

*Factor Decomposition of Wealth Gini Coefficient, 2017*
:::

@@

*Source: Author's calculation based on the 2019 CHFS data.*

*Factor Decomposition of Wealth Gini Coefficient, 2019*
:::

The factor decomposition results for 2017 and 2019 are presented in @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019, respectively. Both years highlight the persistent and dominant role of housing assets in driving wealth inequality in China, contributing over 72% to overall inequality. Specifically, housing assets accounted for 74.2% of total household wealth in 2017 and 73% in 2019, figures notably higher than the 67% reported for 2010 by @tan_earnings_2017 using the same CHFS data. Together with the finding by @knight_why_2022, who documented a substantial increase in housing equity’s contribution from 64% in 2002 to 79% in 2013 based on CHIP data, these results confirm that housing has become an increasingly significant factor in shaping wealth inequality in China during the 21st century.

Financial equity represented another significant factor, with its contribution to overall inequality increasing from approximately 11.2% in 2017 to 13.5% in 2019. Despite its comparatively smaller share of total household wealth, which limits its overall contribution to inequality, financial equity demonstrates a higher degree of internal inequality in its distribution than housing assets. Commercial assets, encompassing agricultural assets, business assets, and storefronts, ranked as the third-largest contributor to overall inequality. This asset class exhibits a very high Gini coefficient, potentially attributable to the concentration of family farming and business operations among a relatively small number of households. Land and vehicles represented smaller contributions to overall inequality. Sharp fluctuations observed in the share and contribution of commercial assets and land to the overall Gini coefficient appear to stem from definitional changes or systemic asset re-evaluation within the survey methodology, warranting caution in interpreting results related to these components across survey waves. Debt generally acts as a negative contributor to overall inequality, suggesting that higher-wealth households tend to hold larger absolute amounts of debt. An exception is noted for other debt, including medical, credit card, education, and miscellaneous debts, which show a slight positive contribution to overall inequality. This indicates that lower-wealth households are more likely to be burdened by consumer debt compared to debt acquired for asset accumulation.

Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019. Housing assets maintained their dominant role, while financial assets became increasingly significant. To explore heterogeneity in wealth inequality patterns, the following sections decompose wealth inequality within specific demographic subgroups.
